### 2020-12-03 R - Deliverable/MSDS 6371 - Project.Rmd
Fix Variable name

---
title: "stats-project"
author: "David Grijalva & Ricco Ferraro"
date: "11/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(olsrr)
library(tidyverse)
library(ggpubr)
library(ggplot2)
library(car)
library(data.table)
library(mltools)
library(olsrr)

# options
#options(scipen = 999)
```

```{r message=FALSE, warning=FALSE}
data_train= read_csv("../Data/train.csv")
data_test = read_csv("../Data/test.csv")
#Convert <chr> to <fctr>
data_train = as.data.frame(unclass(data_train))
data_test = as.data.frame(unclass(data_test))
nrow(data_test)
```

```{r message=FALSE, warning=FALSE}

# Select relevant columns and Neighborhood
train_data_q1 = data_train %>% select(Id, GrLivArea, Neighborhood, SalePrice ) %>% filter(Neighborhood == "NAmes" | Neighborhood ==  "Edwards" | Neighborhood ==  "BrkSide") %>% droplevels()
test_data_q1 = data_test  %>% select(Id, GrLivArea, Neighborhood ) %>% filter(Neighborhood == "NAmes" | Neighborhood ==  "Edwards" | Neighborhood ==  "BrkSide") %>% droplevels()
train_data_q1
train_data_q2 = data_train %>% select(Id, GrLivArea, Neighborhood, SalePrice ) 
train_data_q2$GrLivArea2 = train_data_q2$GrLivArea^2

test_data_q2 = data_test %>% select(Id, GrLivArea, Neighborhood ) 
test_data_q2$GrLivArea2 = test_data_q2$GrLivArea^2
nrow(test_data_q2)
```

```{r}

#lm_fit_experiment = lm(SalePrice ~. -Id-Utilities-Street-PoolQC-MiscFeature-Alley-Fence-FireplaceQu-LotFrontage-GarageCond-GarageFinish-GarageQual-GarageType-GarageYrBlt-BsmtExposure-BsmtFinType2-BsmtCond-BsmtFinType1-BsmtQual-MasVnrArea-MasVnrType-Electrical  , data = train)
#lm_fit_experiment
#Neighborhood_Blmngtn

summary(lm_fit_experiment)
#confint(lm_fit_experiment,  level=0.995)

```

Question of interest
Is the square footage of the living area and the neighborhood  related to the house sales price
neighborhood: NAmes, Edwards and BrkSide 
Note:  For living area provide results in increments of 100 sq. ft.

```{r message=FALSE, warning=FALSE}
# Create scatter plot graphing function

scatter_func = function (data, i){
  title = paste("SalePrice vs GrLivArea by Neighborhood:", i)
  figure = data %>%  filter (Neighborhood==i) %>%  ggplot(aes(x=GrLivArea, y= SalePrice)) + geom_point() + scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) +  geom_smooth(formlula="lm") + ggtitle(paste0(title))
return (figure)
}

hist_boxplot = function (data) {
  g1 = data %>% ggplot() + geom_histogram(aes(x=SalePrice), fill="steelblue") + scale_x_continuous(labels = function(x) format(x, scientific = FALSE)) + ggtitle("Histogram of Sale Price")
  g2 = data %>% ggplot() + geom_histogram(aes(x=GrLivArea), fill="steelblue") + scale_x_continuous(labels = function(x) format(x, scientific = FALSE)) + ggtitle("Histogram of Square Footage of the Living Area")
  g3 = data %>% ggplot() + geom_boxplot(aes(x=Neighborhood, y=SalePrice, fill=Neighborhood)) + scale_y_continuous(labels = function(x) format(x, scientific = FALSE)) + ggtitle("Boxplot of Sale Price by Neighborhood") +  
    theme(axis.text.x = element_text(angle = 60, hjust = 1, size = 6 ))
  figure =  ggarrange(g1, g2, g3, ncol=1)
  return(figure)
}

press = function (model) {
  pr = resid(model)/(1 - lm.influence(model)$hat)
  return(sum(pr^2))
}

```

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
hist_boxplot(train_data_q1)
hist_boxplot(train_data_q2)
```

```{r message=FALSE, warning=FALSE}
for (i in unlist(unique(data_train$Neighborhood)) ) {
  print(scatter_func(data_train, i))
}

```
#	ANALYSIS 1
```{r message=FALSE, warning=FALSE}
one_encode_train_data_q1 =  one_hot(as.data.table(train_data_q1, dropUnusedLevels=TRUE))

#Reference Neighborhood_BrkSide
lm_fit_q1 = lm(SalePrice ~-Id- Neighborhood_BrkSide+ GrLivArea+Neighborhood_Edwards*GrLivArea+Neighborhood_NAmes*GrLivArea, data = one_encode_train_data_q1)

press(lm_fit_q1)
ols_press(lm_fit_q1)

summary(lm_fit_q1)
confint(lm_fit_q1,  level=0.995)

ols_plot_cooksd_bar(lm_fit_q1)
ols_plot_dffits(lm_fit_q1)
ols_plot_diagnostics(lm_fit_q1)

ols_plot_resid_fit(lm_fit_q1)
ols_plot_resid_pot(lm_fit_q1)
ols_pred_rsq(lm_fit_q1)
plot(lm_fit_q1)
```

#	ANALYSIS 2
```{r message=FALSE, warning=FALSE}

one_encode_train_data_q2 =  one_hot(as.data.table(train_data_q2, dropUnusedLevels=TRUE))
one_encode_test_data_q2 =  one_hot(as.data.table(test_data_q2, dropUnusedLevels=TRUE))
#lm_fit_q2 = lm(SalePrice ~. -Id-Neighborhood_BrkSide, data = one_encode_train_data_q2)

#Neighborhood_Blmngtn

lm_fit_q2 = lm(SalePrice ~.-GrLivArea2 -Id-Neighborhood_BrkSide+(Neighborhood_Blmngtn+Neighborhood_Blueste+Neighborhood_BrDale+Neighborhood_ClearCr+Neighborhood_CollgCr+Neighborhood_Crawfor+Neighborhood_Edwards+Neighborhood_Gilbert+Neighborhood_IDOTRR+Neighborhood_MeadowV+Neighborhood_Mitchel+Neighborhood_NAmes+Neighborhood_NPkVill+Neighborhood_NridgHt+Neighborhood_NWAmes+Neighborhood_OldTown+Neighborhood_Sawyer+Neighborhood_SawyerW+Neighborhood_Somerst+Neighborhood_StoneBr+Neighborhood_SWISU+Neighborhood_Timber+Neighborhood_Veenker)*GrLivArea, data = one_encode_train_data_q2)

summary(lm_fit_q2)
confint(lm_fit_q2,  level=0.995)
```
## Model Selection
### Stepwise
```{r message=FALSE, warning=FALSE}
#Stepwise Selection
stepwise_fit = ols_step_both_p(lm_fit_q2, pent =0.05, prem=0.05)
stepwise_fit
```
```{r}
stepwise_fit$predictors
```

```{r}
stepwise_model = stepwise_fit$model
ols_press(stepwise_model)
summary(stepwise_model)
confint(stepwise_model)

```

```{r}
ols_plot_cooksd_bar(stepwise_model)

stepwise_model_outliers = ols_plot_cooksd_bar(stepwise_model, print_plot=F)
stepwise_model_outliers

ols_plot_dffits(stepwise_model)
ols_plot_diagnostics(stepwise_model)

ols_plot_resid_fit(stepwise_model)
ols_plot_resid_pot(stepwise_model)
ols_pred_rsq(stepwise_model)
plot(stepwise_fit)
```

```{r}

stepwise_predictions = c(predict(stepwise_model, newdata = one_encode_test_data_q2))

stepwise_submission = data.frame(Id= one_encode_test_data_q2$Id, SalePrice=stepwise_predictions)
stepwise_submission
#write_csv(stepwise_submission, "../kaggleData/stepwiseSelectionPredictions.csv")
```

```{r message=FALSE, warning=FALSE}
#Forward Selection
fw_fit = ols_step_forward_p(lm_fit_q2, penter=0.05, details = F)
fw_fit

```

```{r}
fw_model = fw_fit$model
ols_press(fw_model)
summary(fw_model)
confint(fw_model)

```

```{r}
ols_plot_cooksd_bar(fw_model)

fw_model_outliers = ols_plot_cooksd_bar(fw_model, print_plot=F)
fw_model_outliers

ols_plot_dffits(fw_model)
ols_plot_diagnostics(fw_model)

ols_plot_resid_fit(fw_model)
ols_plot_resid_pot(fw_model)
ols_pred_rsq(fw_model)
plot(fw_fit)

```

```{r}

fw_predictions = c(predict(fw_model, newdata = one_encode_test_data_q2))

fw_submission = data.frame(Id= one_encode_test_data_q2$Id, SalePrice=fw_predictions)
fw_submission
#write_csv(fw_submission, "../kaggleData/forward SelectionPredictions.csv")
```

```{r message=FALSE, warning=FALSE}
#Backward Selection
bw_fit = ols_step_backward_p(lm_fit_q2, penter=0.05, details = F)
bw_fit
```

```{r}
bw_model = bw_fit$model
ols_press(bw_model)
summary(bw_model)
confint(bw_model)
```

```{r}
ols_plot_cooksd_bar(bw_model)

bw_model_outliers = ols_plot_cooksd_bar(bw_model, print_plot=F)
bw_model_outliers

ols_plot_dffits(bw_model)
ols_plot_diagnostics(bw_model)

ols_plot_resid_fit(bw_model)
ols_plot_resid_pot(bw_model)
ols_pred_rsq(bw_model)
plot(bw_fit)
```
```{r}

bw_predictions = c(predict(bw_model, newdata = one_encode_test_data_q2))
bw_predictions

bw_submission = data.frame(Id= one_encode_test_data_q2$Id, SalePrice=bw_predictions)
bw_submission
#write_csv(bw_submission, "../kaggleData/backwardsSelectionPredictions.csv")
```

```{r message=FALSE, warning=FALSE}
#Backward Selection
custom_fit = ols_step_best_subset(lm_fit_q2,)
custom_fit
```

```{r}
custom_model = custom_fit$model
ols_press(custom_model)
summary(custom_model)
confint(custom_model)
```

```{r}
ols_plot_cooksd_bar(custom_model)

custom_model_outliers = ols_plot_cooksd_bar(custom_model, print_plot=F)
custom_model_outliers

ols_plot_dffits(custom_model)
ols_plot_diagnostics(custom_model)

ols_plot_resid_fit(custom_model)
ols_plot_resid_pot(custom_model)
ols_pred_rsq(custom_model)
plot(custom_fit)
```

```{r}

custom_predictions = c(predict(custom_model, newdata = one_encode_test_data_q2))
custom_predictions

custom_submission = data.frame(Id= one_encode_test_data_q2$Id, SalePrice=custom_predictions)
custom_submission
#write_csv(custom_submission, "../kaggleData/customSelectionPredictions.csv")
```