### 2020-12-03 R - Deliverable/unit14/Untitled.Rmd
Initial Commit

---
title: "R Notebook"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
library(tidyverse)
library(olsrr)
```

```{r}
exclude_df = data.frame(x=c(1,2,3,4),y=c(2,3,3,3))
include_df = data.frame(x=c(1,2,3,4,30),y=c(2,3,3,3,50))
```

````{r}
exclude_fit = lm(y ~x, data=exclude_df)
summary(exclude_fit)
confint(exclude_fit)
ols_plot_cooksd_bar(exclude_fit)
ols_plot_diagnostics(exclude_fit)
```
````{r}
include_fit = lm(y ~x, data=include_df)
summary(include_fit)
confint(include_fit)
ols_plot_cooksd_bar(include_fit)
ols_plot_diagnostics(include_fit)
```
```{r warning=FALSE}
data = read_csv("/Users/dgrijalva/SMU/Classes/Term Fall 2020/Stats-Foundation/MSDS_6371_Stat_Foundations/Unit 14/SAT Data Set.csv")

```
```{r}
fit = lm (Expend ~.-State, data = data)

stepwise_fit = ols_step_both_p(fit, pent =0.05, prem=0.05,  details = F)
stepwise_fit
stepwise_model = stepwise_fit$model
ols_press(stepwise_model)
stepwise_model
summary(stepwise_model)
confint(stepwise_model)

ols_plot_cooksd_bar(stepwise_model)

stepwise_model_outliers = ols_plot_cooksd_bar(stepwise_model, print_plot=F)
stepwise_model_outliers

ols_plot_dffits(stepwise_model)
ols_plot_diagnostics(stepwise_model)

ols_plot_resid_fit(stepwise_model)
ols_plot_resid_pot(stepwise_model)
ols_pred_rsq(stepwise_model)
plot(stepwise_fit)

```